\documentclass[a4,11pt]{memoir}

\setlength{\parindent}{0cm}
\setlength{\parskip}{0.15cm}

\setlrmarginsandblock{2.5cm}{*}{1} 
\setulmarginsandblock{2.5cm}{2.5cm}{*}
\setmarginnotes{2.5mm}{2cm}{1em}
\checkandfixthelayout

\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{
  url,
  fancyvrb,
  multicol
}
\usepackage[colorlinks]{hyperref}
\usepackage{graphicx}
\usepackage[draft]{fixme}
\usepackage{fourier}
% \newcommand\starbreak{\fancybreak{\decosix\quad\decosix\quad\decosix}}

%\newcommand\starbreak{\fancybreak{$*\quad*\quad*$}}

\usepackage[scaled]{berasans}



% \usepackage[T1]{fontenc}
% \usepackage{tgpagella}
% \usepackage[sc]{mathpazo}
\usepackage{authblk}
% \usepackage{times}
\usepackage{pldoc2}


%% Margin notes
% \setlength{\marginparwidth}{3cm}
\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
{\raggedright\footnotesize \fbox{\textbf{#1}}}}
%%%%%%%%%%%


\sloppy
\makeindex

\title{kLog\\
  User's Manual}
\author[1,2]{Paolo Frasconi}
\author[2]{Fabrizio Costa}
\author[2]{Luc De Raedt}
\author[2]{Kurt De Grave}
\affil[1]{Dipartimento di Sistemi e Informatica, Universit{\`a} degli
  Studi di Firenze, via di Santa Marta 3, I-50139 Firenze, Italy}
\affil[2]{Departement Computerwetenschappen, Katholieke Universiteit
  Leuven, Celestijnenlaan 200A B-3001 Heverlee, Belgium}

% \setcounter{secnumdepth}{3}% to get numbered subsections
\setsecnumdepth{subsection}% to get numbered subsections

\begin{document}

\maketitle
\chapterstyle{demo3}

\frontmatter

\tableofcontents
% \setlength{\unitlength}{1pt}
% \clearpage
% \listoffigures
% \clearpage
% \listoftables
% \clearpage
% \listofegresults

\mainmatter
\chapterstyle{ell}


\chapter{Introduction}

This document provides basic information for kLog, a logical and
relational language for developing kernel-based learning systems.
kLog has been developed at K.U. Leuven and at Università degli Studi
di Firenze by P. Frasconi, F. Costa, L. De Raedt and K. De Grave.

License?

\chapter{Tutorial}
In this tutorial we explain by examples how to use kLog in several
relational learning data sets. You are encouraged to scan the whole
tutorial since most concepts are only explained the first time they
are used. 

kLog learns from interpretations, i.e. training data consist of sets
of ground atoms that are true in a given interpretation. An
alternative view is that each interpretation is an instance of a
relational database.

\section{Bursi}
This is a small molecules data set and the goal is the discrimination
between mutagenic and non-mutagenic compounds. Every molecule
corresponds to one interpretation so the kLog job is binary
classification of interpretations.

\marginpar{kLog data sets}
Data sets in kLog are stored as Prolog files consisting of collections
of ground facts. The predefined predicate \predref{interpretation}{2}
\index{interpretation/2} takes as its first argument an interpretation
identifier and as its second argument a ground term.

In the example below \predref{a,2} describes atoms, \predref{b,3}
describes chemical bonds, \predref{sub,3} functional groups,
\predref{subat,3} atom membership in functional groups,
\predref{target,1} describes the category of the molecule
(\texttt{mutagen} is the positive class, \texttt{nonmutagen} the
negative class), and \predref{linked,2} describes how functional
groups are linked together.

\begin{linumcode}
interpretation(i788,a(a1,c)).
interpretation(i788,a(a2,c)).
interpretation(i788,a(a3,cl)).
interpretation(i788,a(a4,br)).
interpretation(i788,a(a5,n)).
interpretation(i788,a(a6,h)).
interpretation(i788,b(a1,a2,1)).
interpretation(i788,b(a1,a3,1)).
interpretation(i788,b(a1,a4,1)).
interpretation(i788,b(a2,a5,3)).
interpretation(i788,b(a1,a6,1)).
interpretation(i788,sub(f1,halide,1)).
interpretation(i788,sub(f2,halide,1)).
interpretation(i788,sub(f3,nitrile,2)).
interpretation(i788,sub(f4,aliphatic_chain,1)).
interpretation(i788,subat(f1,a3,1)).
interpretation(i788,subat(f2,a4,1)).
interpretation(i788,subat(f3,a2,1)).
interpretation(i788,subat(f3,a5,2)).
interpretation(i788,subat(f4,a1,1)).
interpretation(i788,target(mutagen)).
interpretation(i788,linked(f4,[link(f3,a2,a1),link(f1,a3,a1),link(f2,a4,a1)],[branch(1,3)],saturated)).
\end{linumcode}

\begin{linumcode}
begin_domain.
signature atm(atom_id::self, element::property)::intensional.
atm(Atom, Element) :-
    a(Atom,Element), \+(Element=h).

signature bnd(atom_1@b::atm, atom_1@b::atm, type::property)::intensional.
bnd(Atom1,Atom2,Type) :-
    b(Atom1,Atom2,NType), describeBondType(NType,Type), atm(Atom1,_), atm(Atom2,_).

signature fgroup(fgroup_id::self, group_type::property)::intensional.
fgroup(Fg,Type) :- sub(Fg,Type,_).

signature fgmember(fg::fgroup, atom::atm)::intensional.
fgmember(Fg,Atom):- subat(Fg,Atom,_), atm(Atom,_).

signature fg_fused(fg1@nil::fgroup, fg2@nil::fgroup, nrAtoms::property)::intensional.
fg_fused(Fg1,Fg2,NrAtoms):- fus(Fg1,Fg2,NrAtoms,_AtomList).

signature fg_connected(fg1@nil::fgroup, fg2@nil::fgroup, bondType::property)::intensional.
fg_connected(Fg1,Fg2,BondType):-
    con(Fg1,Fg2,Type,_AtomList),describeBondType(Type,BondType).

signature fg_linked(fg::fgroup, alichain::fgroup, saturation::property)::intensional.
fg_linked(FG,AliChain,Sat) :-
    linked(AliChain,Links,_BranchesEnds,Saturation),
    (Saturation = saturated ->
     Sat = saturated
    ;
     Sat = unsaturated
    ),
    member(link(FG,_A1,_A2),Links).

signature mutagenic::intensional.
mutagenic :- target(mutagen).
end_domain.
\end{linumcode}

\section{Biodeg}
\section{UW-CSE}
\section{WebKB}
\section{IMDB}

% The rest of this LaTeX document is generated using the LaTeX
% backend of PlDoc, The SWI-Prolog documentation system

\chapter{Main components of the kLog Prolog library}
\input{syntax}
\input{db}
\input{graphicalize}
\input{kfold}
\input{learn}
\input{flags}
\input{repair}

\chapter{Advanced components}
\input{timing}
\input{goals}
\input{utils}
\input{cinterface}

\backmatter
\chapterstyle{demo3}

\clearpage
\twocolindex
%\renewcommand{\chaptermark}[1]{}
\renewcommand{\preindexhook}{%
The first page number is usually, but not always, the primary reference to
the indexed topic.\vskip\onelineskip}
\indexintoc

%%%\raggedright  does nasty things to index entries
\printindex
\end{document}
